---
title: "FNA transcriptomics"
author: "Ed Carr, Linterman lab"
date: "8 Dec 2020"
output: 
  pdf_document: 
editor_options: 
  chunk_output_type: console
---

```{r echo=F}
Sys.setenv(PATH=paste(Sys.getenv("PATH"),
                      "/bi/home/carre/texlive/2017/bin/x86_64-linux/",sep=":"))

```

```{r setup, include=FALSE}

## set a few options for pdf formatting.
knitr::opts_chunk$set(echo = F, tidy = T, warning = F, message = F, fig.width = 7, fig.height = 9.5, tidy.opts=list(width.cutoff=60))
```

# Example code

Example code collected from orignal bash scripts, are shown for single samples.

cellranger v 5.0.0.

Reference files from 10x Genomics:
- refdata-cellranger-GRCh38-1.2.0 (transcriptome)
- refdata-cellranger-vdj-GRCh38-alts-ensembl-5.0.0 (vdj)

Seurat import onwards in later Rmd files.

## 1. Download the scRNAseq and Ig BCR libraries from SRA

```{bash example_code_to_download_from_SRA, echo = TRUE, eval = F}

## From http://SRAexplorer.info

#!/bin/bash
curl -L ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR112/056/SRR11233656/SRR11233656_1.fastq.gz -o SRR11233656_scRNA-seq_of_Homo_sapiens_adult_male_FNA_from_lymph_node_d0_1.fastq.gz &
curl -L ftp://ftp.sra.ebi.ac.uk/vol1/fastq/SRR112/056/SRR11233656/SRR11233656_2.fastq.gz -o SRR11233656_scRNA-seq_of_Homo_sapiens_adult_male_FNA_from_lymph_node_d0_2.fastq.gz

## All transcriptomic fq moved into a directory called 'fq_dir'

## All vdj fq moved into a directory called 'vdj_fqdir'.
```

## 2. Transcriptomic fq counting

```{bash cellranger_counting, echo=TRUE, eval=FALSE}

module load cellranger

cellranger count --id TurnerSRR11233645 --fastqs fq_dir --sample SRR11233645_scRNA-seq_of_Homo_sapiens_adult_male_FNA_from_lymph_node_d60 --transcriptome /bi/scratch/Genomes/10x/human/refdata-cellranger-GRCh38-1.2.0


```

## 3. VDJ fq cellranger

```{bash cellranger_vdj, echo = TRUE, eval = FALSE}

module load cellranger

cellranger vdj --id Turner_VDJ_SRR11233622 --fastqs vdj_fqdir --sample SRR11233622_single_cell_VDJ_of_Homo_sapiens_adult_male_FNA_from_lymph_node_d60 --reference /bi/home/carre/refdata-cellranger-vdj-GRCh38-alts-ensembl-5.0.0

```


# SessionInfo

```{r sessionInfo}

sessionInfo()

```