---
title: ""
author: ""
date: ""
output: pdf_document
header-includes:
    - \pagenumbering{gobble}
editor_options: 
  chunk_output_type: console
---

```{r echo=F}
Sys.setenv(PATH=paste(Sys.getenv("PATH"),
                      "/bi/home/carre/texlive/2017/bin/x86_64-linux/",sep=":"))

```

```{r setup, include=FALSE}

## set a few options for pdf formatting.
knitr::opts_chunk$set(echo = F, tidy = T, warning = F, message = F, fig.width = 7, fig.height = 9, tidy.opts=list(width.cutoff=60))
```

```{r data_load}



library(tidyverse)
library(ggpubr) # for themes for plots.


load(file = "../cohort_2016_17/Concatenated_index_master_fcs_gated_in_FlowJo_2_table_output.RData")

merged.data <- as_tibble(merged.data)

```

```{r first_panel_is_flow_gating}

# See flowsinglecell repository for details.
## flowWorkspace used to parse a flowjo .wsp, with gated files.
## ggcyto used to plot, with the gpubr::theme_pubr() super-imposed.
## arranged as a single row of plots and the title section removed.

load(file = "../cohort_2016_17/Example_of_sc_sort_gating_from_665X_d42.RData")




```

```{r plots, fig.height=9}
##
# Plots
# The plots need some adjustments to labels etc.
# These are done in the figure.Rmd in B_flu_scRNAseq repository.
###
## choose a y
## (then the next few lines do not need adjusting)
merged.data$y1 <- merged.data$`Lymphocytes/Single Cells/Single Cells/DUMP-/CD19+SA-/hA_APC+hA_PE+ | Freq. of Parent (%)`

y1 <- "y1"
ylab1 <- "\nHA+ B cells \n[% of live B cells]"

y <- "num_HA_B_cells_per_10e6_PBMC"
ylab <- "<br>number of HA B cells per 10<sup>7</sup> PBMC"
##

# Reshape for a paired plot
data.for.ggpaired <- reshape2::dcast(
  merged.data, PID+age ~ day, value.var = y)

data.for.ggpaired1 <- reshape2::dcast(
  merged.data, PID+age ~ day, value.var = y1)


cowplot::plot_grid(

  
  cowplot::plot_grid(p2, labels = "A"),
  
  cowplot::plot_grid(
  # HA positive B cells [% of B cells] days 0 vs 42
  # Facet by day, to compare the day 42 %s - no difference.
  
  ggstripchart(merged.data, x = "age", 
               y = y1,
               facet.by = "day",
               ylab = ylab1) +
    geom_boxplot(outlier.shape = NA, fill = NA) + 
    stat_compare_means(method = "wilcox",
                       label.y = 0.35,
                       paired = FALSE,
                       comparisons = list(c(1,2))) +
    rotate_x_text() + 
    theme(strip.background = element_rect(fill = NA)),
  
  ### ROW 2
  
  ggpaired(data.for.ggpaired, cond1 = "d0", cond2 = "d42", facet.by = "age",
         ylab = ylab, line.color = "grey",
         xlab = "day") +
  stat_compare_means(method = "wilcox",
                     label.y = 3000,
                     paired = TRUE,
                     comparisons = list(c(1,2))) +
    theme(strip.background = element_rect(fill = NA),
          axis.title.y = ggtext::element_markdown()),





ggstripchart(merged.data, x = "age", 
          y = y,
          facet.by = "day",
          ylab = ylab) +
  geom_boxplot(outlier.shape = NA, fill = NA) + 
  stat_compare_means(method = "wilcox",
                     label.y = 3000,
                     paired = FALSE,
                     comparisons = list(c(1,2))) +
  rotate_x_text() +
  theme(strip.background = element_rect(fill = NA),
          axis.title.y = ggtext::element_markdown()),



ncol = 3, labels = LETTERS[2:4]),

ncol = 1, rel_heights = c(5,4))

###
# Fold expansion
data.for.ggpaired$expansion <- data.for.ggpaired$d42/data.for.ggpaired$d0

# data.for.ggpaired %>%
#   group_by(age) %>%
#   dplyr::summarise(median = median(expansion))

###
# Useful to calculate the median cell counts for figure legends:
tab <- merged.data %>% 
  group_by(day, age) %>%
  summarise(m = round(median(num_HA_B_cells_per_10e6_PBMC)))  

```


\newpage
__Supplementary figure 1: Single cell sorting strategy for haemagglutinin specific B cells and abundance of haemagglutinin specific B cells__

(A) Example flow sorting gating strategy on B cells. B cells were negatively separated from PBMC using magnetic sorting prior to flow sorting.
(B) The proportion of haemagglutinin binding B cells (as % of live B cells), is not significantly different between age groups either before, or 6 weeks after TIV immunization.
(C-D) The number of haemagglutinin binding B cells per 10^7^ PBMC analysed by age and days post vaccination.
(D) The number of haemagglutinin binding B cells is higher in 22-36 year olds at day 42. The number of haemagglutinin binding B cells was calculated as in (C).
In (B) and (D), the _P_ values shown are from an unpaired two-tailed Mann-Whitney test. In (C), samples from the same individual are indicated with a grey line. The _P_ values shown are from a paired two-tailed Mann-Whitney test.


\newpage

# SessionInfo

```{r sessionInfo}

sessionInfo()

```