---
title: ""
author: ""
date: ""
output:
  pdf_document:
    keep_tex: no
header-includes:
    - \pagenumbering{gobble}
editor_options: 
  chunk_output_type: console
---

```{r echo=F}
# Sys.setenv(PATH=paste(Sys.getenv("PATH"),
#                       "/bi/home/carre/texlive/2017/bin/x86_64-linux/",sep=":"))

```

```{r setup, include=FALSE}

## set a few options for pdf formatting.
knitr::opts_chunk$set(echo = F, tidy = T, warning = F, message = F, fig.width = 7, fig.height = 9, tidy.opts=list(width.cutoff=60))
```

```{r data_load}

library(MultiAssayExperiment)
library(SingleCellExperiment)
library(tidyverse)
library(ggpubr) # for themes for plots.


load(file = "../cohort_2016_17/Concatenated_index_master_fcs_gated_in_FlowJo_2_table_output.RData")
load(file = "../cohort_2016_17/data/SCE_QC_pass_finalised.RData")
merged.data <- as_tibble(merged.data)

```

```{r first_panel_is_flow_gating}

# See flowsinglecell repository for details.
## flowWorkspace used to parse a flowjo .wsp, with gated files.
## ggcyto used to plot, with the gpubr::theme_pubr() super-imposed.
## arranged as a single row of plots and the title section removed.

load(file = "../cohort_2016_17/Example_of_sc_sort_gating_from_665X_d42.RData")




```

```{r hai}
load(file = "../cohort_2016_17/ALFNA16_MultiAssayExperiment.RData")
hai.data <- wideFormat(ALFNA16[ , , 
                                # select assays with names that match:
                                # 'HAI'
                                # which returns all data from these two assays at all available days.
                                grepl(names(assays(ALFNA16)), pattern = "HAI") ],
                      # Return ages as well:
                       colDataCols = "Age.Group")


hai.data <- as.data.frame(hai.data)

hai.data <- hai.data %>% filter(primary %in% sce$PID)
```

```{r plots, fig.height=9}
##
# Plots
# The plots need some adjustments to labels etc.
# These are done in the figure.Rmd in B_flu_scRNAseq repository.
###
## choose a y
## (then the next few lines do not need adjusting)
merged.data$y1 <- merged.data$`Lymphocytes/Single Cells/Single Cells/DUMP-/CD19+SA-/hA_APC+hA_PE+ | Freq. of Parent (%)`

y1 <- "y1"
ylab1 <- "\nHA+ B cells \n[% of live B cells]"

y <- "num_HA_B_cells_per_10e6_PBMC"
ylab <- "<br>number of HA B cells per 10<sup>7</sup> PBMC"
##

# Reshape for a paired plot
data.for.ggpaired <- reshape2::dcast(
  merged.data, PID+age ~ day, value.var = y)

data.for.ggpaired1 <- reshape2::dcast(
  merged.data, PID+age ~ day, value.var = y1)


cowplot::plot_grid(
  
  cowplot::plot_grid(
  ggplot(data = 
               hai.data %>%
               dplyr::select(primary, Age.Group, 
                             contains("HAI_d"), -contains("log2")) %>%
         mutate(Age.Group=str_replace_all(Age.Group, c("young"="18-22yo", "old"="67-86yo"))) %>%
               mutate("Age" = factor(paste0(Age.Group))) %>%
               dplyr::select(-Age.Group) %>%
               set_colnames(
                 gsub(colnames(.), 
                      pattern = "HAI_d|_titre", replacement = "")) %>%
               reshape2::melt( id_vars = c(primary, Age)),
       aes(x = Age, y=value,fill = Age, group=primary)) +
         geom_violin(aes(fill=NULL), col="grey", trim=F) +
  geom_point(shape =21, position = position_dodge(width=0.4)) +
  scale_fill_manual(values = c("white", "darkgrey")) +
  labs(x ="age", y="\nA/Cal09 HAI titer") +
  stat_compare_means(method="wilcox", paired=F, comparisons=list(c(1,2))) +
  scale_y_continuous(trans = "log2", limits = c(1, 10000)) +

  facet_grid(.~variable) +
  theme_pubr(legend = "none") +
  theme(strip.background = element_rect(fill = "NA")), labels = "A"),

  
  cowplot::plot_grid(p2, labels = "B"),
  
  cowplot::plot_grid(
  # HA positive B cells [% of B cells] days 0 vs 42
  # Facet by day, to compare the day 42 %s - no difference.
  
  ggstripchart(merged.data, x = "age", 
               y = y1,
               facet.by = "day",
               ylab = ylab1) +
    geom_boxplot(outlier.shape = NA, fill = NA) + 
    stat_compare_means(method = "wilcox",
                       label.y = 0.35,
                       paired = FALSE,
                       comparisons = list(c(1,2))) +
    rotate_x_text() + 
    theme(strip.background = element_rect(fill = NA)),
  
  ### ROW 2
  
  ggpaired(data.for.ggpaired, cond1 = "d0", cond2 = "d42", facet.by = "age",
         ylab = ylab, line.color = "grey",
         xlab = "day") +
  stat_compare_means(method = "wilcox",
                     label.y = 3000,
                     paired = TRUE,
                     comparisons = list(c(1,2))) +
    theme(strip.background = element_rect(fill = NA),
          axis.title.y = ggtext::element_markdown()),





ggstripchart(merged.data, x = "age", 
          y = y,
          facet.by = "day",
          ylab = ylab) +
  geom_boxplot(outlier.shape = NA, fill = NA) + 
  stat_compare_means(method = "wilcox",
                     label.y = 3000,
                     paired = FALSE,
                     comparisons = list(c(1,2))) +
  rotate_x_text() +
  theme(strip.background = element_rect(fill = NA),
          axis.title.y = ggtext::element_markdown()),



ncol = 3, labels = LETTERS[3:5]),

ncol = 1, rel_heights = c(2,5,4))

###
# Fold expansion
data.for.ggpaired$expansion <- data.for.ggpaired$d42/data.for.ggpaired$d0

# data.for.ggpaired %>%
#   group_by(age) %>%
#   dplyr::summarise(median = median(expansion))

###
# Useful to calculate the median cell counts for figure legends:
tab <- merged.data %>% 
  group_by(day, age) %>%
  summarise(m = round(median(num_HA_B_cells_per_10e6_PBMC)))  

```


\newpage
__Supplementary figure 1: Single cell sorting strategy for haemagglutinin specific B cells and abundance of haemagglutinin specific B cells__

(B) Example flow sorting gating strategy on B cells. B cells were negatively separated from PBMC using magnetic sorting prior to flow sorting.
(C) The proportion of haemagglutinin binding B cells (as % of live B cells), is not significantly different between age groups either before, or 6 weeks after TIV immunization.
(D-E) The number of haemagglutinin binding B cells per 10^7^ PBMC analysed by age and days post vaccination.
(E) The number of haemagglutinin binding B cells is higher in 22-36 year olds at day 42. The number of haemagglutinin binding B cells was calculated as in (C).
In (C) and (E, the _P_ values shown are from an unpaired two-tailed Mann-Whitney test. In (D), samples from the same individual are indicated with a grey line. The _P_ values shown are from a paired two-tailed Mann-Whitney test.


\newpage

# SessionInfo

```{r sessionInfo}

sessionInfo()

```