---
title: "SCE assembly for the 2016-17 cohort"
author: "EJC"
date: "25/06/2020, based on code from May 2019"
output: 
  pdf_document: 
    keep_tex: no
editor_options: 
  chunk_output_type: console
---

# Setup

The Babraham compute cluster does not contain a global tex installation, so a local tex is added to $PATH to allow knitting to pdf.

```{r}
Sys.setenv(PATH=paste(Sys.getenv("PATH"),
                      "/bi/home/carre/texlive/2017/bin/x86_64-linux/",sep=":"))

```

```{r setup, include=FALSE}

## set a few options for pdf formatting.
knitr::opts_chunk$set(echo = T, tidy = T, warning = F, message = T, fig.width = 7, fig.height = 4.5, tidy.opts=list(width.cutoff=60))
```

# Load QC-pass SCE

```{r get_sce}
load(file = "data/SCE_QC_pass.RData")
```

# Normalise by deconvolution, PCA, UMAP, find PB cluster and remove

For convenience PCA + UMAP on the final SCE is re-run (to avoid needing to re-run it for each figure.Rmd).

```{r}
library(SingleCellExperiment)
library(scran)
library(scater)

# Normalisation by deconvolution

set.seed(100)
clust <- quickCluster(sce) 
# table(clust)
deconv.sf <- calculateSumFactors(sce, cluster=clust)
# summary(deconv.sf)

sce <- computeSumFactors(sce, cluster=clust, min.mean=0.1)
sce <- logNormCounts(sce)

# Select 25% of genes with largest variance
dec <- modelGeneVar(sce)
hvg <- getTopHVGs(dec, prop=0.25)

set.seed(10000)
sce <- runPCA(sce, ncomponents=50, subset_row=hvg)
set.seed(100000)
sce <- runUMAP(sce, dimred = 'PCA', external_neighbors=TRUE)

# Clustering.
set.seed(1000000)
g <- buildSNNGraph(sce, use.dimred = 'PCA')
set.seed(10000000)
sce$clusters <- factor(igraph::cluster_louvain(g)$membership)


# So there is always 1 cluster whose row has 9 zeros in
# Let's select that programmatically.
cluster.selector <- table(Cluster=sce$clusters, Batch=sce$library)
cluster.to.discard <- c(1:nlevels(sce$clusters))[apply(cluster.selector, 1, function(x) sum(x == 0)) == 9]

# Discard the PB cluster, then re-do analysis:
## The PB cluster is so different, that % hvg or # PCs does not influence their clustering apart
sce.old <- sce
```

```{r}
for(p in c(0.1,0.15, 0.2, 0.25, 0.3)){
  load(file = "data/SCE_QC_pass.RData")
sce <- sce[, ! sce.old$clusters == cluster.to.discard]

set.seed(100)
clust <- quickCluster(sce) 
# table(clust)
deconv.sf <- calculateSumFactors(sce, cluster=clust)
# summary(deconv.sf)

sce <- computeSumFactors(sce, cluster=clust, min.mean=0.1)
sce <- logNormCounts(sce)

# Select 25% of genes with largest variance
dec <- modelGeneVar(sce)

hvg <- getTopHVGs(dec, prop=p)

for(npcs in c(15, 20, 25, 30, 35, 40, 45, 50, 55, 60)){

set.seed(10000)
  sce <- runPCA(sce, ncomponents=npcs, subset_row=hvg)
set.seed(100000)
sce <- runUMAP(sce, dimred = 'PCA', external_neighbors=TRUE)

# Clustering.
set.seed(1000000)
g <- buildSNNGraph(sce, use.dimred = 'PCA')
set.seed(10000000)
sce$clusters <- factor(igraph::cluster_louvain(g)$membership)

library(ggpubr)

plots <- cowplot::plot_grid(
plotUMAP(sce, colour_by="clusters") + 
     theme_pubr(legend = "right") + 
     guides(fill = guide_legend(title = "Louvain\ncluster",
                                title.position = "top")) + 
     theme(legend.key.size = unit(0, 'lines'),
           legend.margin = margin(0,0,0,0, 'lines'),
           aspect.ratio = 1) + rotate_x_text() +
  ggtitle(paste("HVG", p, "%;", npcs, "PCAs")),

cowplot::plot_grid(
  # UMAP both day 0 and day 42
  # Colours from index flow IgD
  plotUMAP(sce, colour_by="IgD.BUV737") +
    viridis::scale_fill_viridis(option="inferno") + 
    labs(fill = "Surface\nIgD") +
    theme_pubr() + 
    theme(legend.position = "right",
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          legend.key.width = unit(6,"points"),
          legend.key.height = unit(12,"points"),
          legend.margin = margin(0,0,0,0, 'lines'),
          aspect.ratio = 1),
  plotUMAP(sce, colour_by="CD27.BV711") + 
    viridis::scale_fill_viridis(option="inferno") +
     labs(fill = "Surface\nCD27") +
     theme_pubr() +
     theme(legend.position = "right",
           axis.ticks = element_blank(),
           axis.text = element_blank(),
           legend.key.width = unit(6,"points"),
           legend.key.height = unit(12,"points"),
           legend.margin = margin(0,0,0,0, 'lines'),
           aspect.ratio = 1),
  
  # UMAP both day 0 and day 42
  # Colours from index flow CD21
  plotUMAP(sce, colour_by="CD21.PE.cy7") +
    viridis::scale_fill_viridis(option="inferno") +
    labs(fill = "Surface\nCD21") +
    theme_pubr() +
    theme(legend.position = "right",
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          legend.key.width = unit(6,"points"),
          legend.key.height = unit(12,"points"),
          legend.margin = margin(0,0,0,0, 'lines'),
          aspect.ratio = 1),
  
  # UMAP both day 0 and day 42
  # Colours from index flow CD28
  plotUMAP(sce, colour_by="CD38.BV421") + 
    viridis::scale_fill_viridis(option="inferno") +
    labs(fill = "Surface\nCD38") +
    theme_pubr() +
    theme(legend.position = "right",
          axis.ticks = element_blank(),
          axis.text = element_blank(),
          legend.key.width = unit(6,"points"),
          legend.key.height = unit(12,"points"),
          legend.margin = margin(0,0,0,0, 'lines'),
          aspect.ratio = 1),
  
  ncol = 2),
ncol = 2)

print(plots)
}}
```

# Save final SCE: NTC removed, QC passing cells, plasmablasts removed, with dimensionalty reduction pre-calculated

```{r}

# save(sce, file = "data/SCE_QC_pass_finalised.RData")

```
# SessionInfo

```{r sessionInfo}

sessionInfo()

```